---
title: "Lab 6 Penguins"
author: "Angie Bouche"
date: "November 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in the penguins.csv data

Packages: tidyverse, car, onewaytests

```{r}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read.csv("penguins.csv")

```

Our quesiton: Is there a significant difference in mean flipper lengths for female chinstrap. gentoo, and adelie penguins near palmer archipelago?

Factor- one, species
Level- three, each species
Variable- one continuous, flipper length

Maybe we can use a one-way ANOVA> Assumptions:
-Independant observations
-Equal variance (Levernes Test)
-Normality (Histogram, QQ)

```{r}
counts <- with(penguins, table(species, sex)) #create a table for number of each sex by species

penguins2 <- penguins %>% 
  mutate(
    common_name = 
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")


```

Exploratory graphs and Levene's test for equal variances

```{r}
hists <- ggplot(penguins2, aes(x= flipper_length))+
  geom_histogram (aes(fill=common_name))+
  facet_wrap(~common_name, scale = "free")

hists

qq<- ggplot(penguins2, aes(sample= flipper_length))+
  geom_qq (aes(color = common_name))+
  facet_wrap(~common_name, scale = "free")

qq

####Based on QQ plots and historgrams, observations for each group appear normally distributed. Also n >30 for all groups, so means will bbe normally distributed by CLT.


```

Evaluate variances and perfirn a Levenes Test for equal variances

```{r}
#If the largest saple variance is <4x greater than the smallest sample vairne, then these are close enough to consider variances equal.

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

variances #The largest variance (33.1) is definitely not more than 4x grater than the smallest variance (15.2).

#leveneTest in the 'car' package

levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins2)
levene_penguin

#H0: No difference in variances (variances are equal)
#HA: Variances are NOT equal

#There is no significant differnce in variances accross species.

```

One-Way ANOA for penguin flipper lengths by species (common_name)

```{r}
#H0: Means across all groups are equal
#HA: At least two group means differ significantly

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)

summary(penguin_aov)

#Post hoc testing by Tukey's HSD
#H0 no significant difference in pairwise means
#HA significant difference in pairwise means

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

#aov.test function in the 'onewaytests' package DOES allow for inline referencing

```

Mean flipper lengths of female penguins differed significantly for all species studied (one-way anova, F(df) = statistic, p <0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05)