---
title: "Lab 6 Penguins"
author: "Angie Bouche"
date: "November 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in the penguins.csv data

Packages: tidyverse, car, onewaytests

```{r}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read.csv("penguins.csv")

```

Our quesiton: Is there a significant difference in mean flipper lengths for female chinstrap. gentoo, and adelie penguins near palmer archipelago?

Factor- one, species
Level- three, each species
Variable- one continuous, flipper length

Maybe we can use a one-way ANOVA> Assumptions:
-Independant observations
-Equal variance (Levernes Test)
-Normality (Histogram, QQ)

```{r}
counts <- with(penguins, table(species, sex)) #create a table for number of each sex by species

penguins2 <- penguins %>% 
  mutate(
    common_name = 
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")


```


